---
title: "Lab 04 - La Quinta is Spanish for next to Denny's, Pt. 1"
author: "Adam Paul"
date: "2/11"
output: github_document
---

### Load packages and data


```{r load-packages, message = FALSE}
library(tidyverse) 
```

```{r load-data, message = FALSE}
states <- read_csv("data/states.csv")
load("data/dennys.rda")
load("data/laquinta.rda")
```

>Getting this to work took a lot longer than I wanted to, until I finally realized the problem was that I hadn't called the data within my markdown. It was running but not knitting.

### Exercise 1

1. What are the dimensions of the `Denny’s` dataset? (Hint: Use inline R code and functions like `nrow` and `ncol` to compose your answer.) What does each row in the dataset represent? What are the variables?

bind_rows(dennys)
```{r }
glimpse(dennys)
```
There are 1,643 rows, which are each an entry of denny's locations. There are six columns, indicating "address", "city", "state", "zip-code", "longitude", and "latitude."

```{r }
glimpse(laquinta)
```
There are 909 rows, indicating each listed La Quinta and its associated variable responses. There are six columns, all the same as Denny's; indicating "address", "city", "state", "zip-code", "longitude", and "latitude."

### Exercise 2

Remove this text, and add your answer for Exercise 2 here.
Add code chunks as needed.
Don't forget to label your code chunk.
Do not use spaces in code chunk labels.

### Exercise 3

...

### Exercise 4

...

### Exercise 5

...

### Exercise 6

...

 
Add exercise headings as needed.

