---
title: "Lab 04 - La Quinta is Spanish for next to Denny's, Pt. 1"
author: "Adam Paul"
date: "2/11"
output: github_document
---

### Load packages and data


```{r load-packages, message = FALSE}
library(tidyverse) 
```

```{r load-data, message = FALSE}
states <- read_csv("data/states.csv")
load("data/dennys.rda")
load("data/laquinta.rda")
```

>Getting this to work took a lot longer than I wanted to, until I finally realized the problem was that I hadn't called the data within my markdown. It was running but not knitting.

### Exercise 1

*1. What are the dimensions of the `Denny’s` dataset? (Hint: Use inline R code and functions like `nrow` and `ncol` to compose your answer.) What does each row in the dataset represent? What are the variables?*

```{r dennys dimensions}
glimpse(dennys)
```

There are 1,643 rows, which are each an entry of denny's locations. There are six columns, indicating "address", "city", "state", "zip-code", "longitude", and "latitude."

*2. What are the dimensions of the La Quinta’s dataset? What does each row in the dataset represent? What are the variables?*

```{r laquinta dimensions}
glimpse(laquinta)
```
There are 909 rows, indicating each listed La Quinta and its associated variable responses. There are six columns, all the same as Denny's; indicating "address", "city", "state", "zip-code", "longitude", and "latitude."

We would like to limit our analysis to Denny’s and La Quinta locations in the United States.

*3. Take a look at the websites that the data come from. Are there any La Quinta’s locations outside of the US? If so, which countries? What about Denny’s?*

There are both La Quinta's in Mexico, Canada, China, New Zealand, Honduras, Turkey, UAE, Chile, and Columbia.

There are actually Denny's in Canada, Mexico, Costa Rica, Chile, Honduras, El Salvador, the United Kingdom, New Zealand, according to Google.

*4. Now take a look at the data. What would be some ways of determining whether or not either establishment has any locations outside the US using just the data (and not the websites). Don’t worry about whether you know how to implement this, just brainstorm some ideas. Write down at least one as your answer, but you’re welcomed to write down a few options too.*

One way would be to look for `state` responses that are not U.S. states, likely returning as NA. Alternatively, if you hated your free time, you could do the same by accounting for all possible U.S. Zip codes. Likewise, if you hated your free time even more you could account for the borders of the U.S. by `latitude` and `longitude` and look for responses that fail to fall within those criteria.

*5. Find the Denny’s locations that are outside the US, if any. To do so, filter the Denny’s locations for observations where `state` is not in `states$abbreviation`. The code for this is given below. Note that the `%in%` operator matches the states listed in the state variable to those listed in `states$abbreviation`. The `!` operator means not. Are there any Denny’s locations outside the US?*

```{r filtering for non-US Dennys}
dennys %>%
  filter(!(state %in% states$abbreviation))
```

This turned up no Denny's outside the U.S.

*6. Add a country variable to the Denny’s dataset and set all observations equal to `"United States"`. Remember, you can use the `mutate` function for adding a variable. Make sure to save the result of this as dennys again so that the stored data frame contains the new variable going forward.*

```{r dennys newvar United States}
dennys <- dennys %>%
  mutate(country = "United States")
```

*7. Find the La Quinta locations that are outside the US, and figure out which country they are in. This might require some googling. Take notes, you will need to use this information in the next exercise.*

```{r la quinta outside USA}
laquinta %>%
  filter(!(state %in% states$abbreviation)) %>%
    view()
```

There are 11 in Mexico, 2 in Canada, and 1 in Colombia.

*8. Add a country variable to the La Quinta dataset. Use the case_when function to populate this variable. You’ll need to refer to your notes from Exercise 7 about which country the non-US locations are in.*

```{r laquina newvar United States}
laquinta <- laquinta %>%
  mutate(country = case_when(
    state %in% state.abb ~ "United States",
    state %in% c("ON", "BC") ~ "Canada",
    state == "ANT"           ~ "Colombia",
    state %in% c("AG", "QR", "CH", "NL", "VE", "PU", "SL", "FM") ~ "Mexico"
  ))
```

Going forward we will work with the data from the `United States` only. All Denny’s locations are in the United States, so we don’t need to worry about them. However we do need to filter the La Quinta dataset for locations in United States.

```{r laquinta usa filter}
laquinta <- laquinta %>%
  filter(country == "United States")
```

*9. Which states have the most and fewest Denny’s locations? What about La Quinta? Is this surprising? Why or why not?*

```{r state frequency}
sort(table(dennys$state))
sort(table(laquinta$state))
```

The state with the most Denny's is California with 403, and the least is Delaware with 1.

The state with the most La Quinta locations is Texas with 237, and the least is Maine with one. 

Neither of these is surprising, as Denny's started on the west coast and La Quinta is part of a megacorporation that would use their different names based on location, and a Spanish name makes the most sense in Texas. Also, they started in Texas.


*10. Which states have the most Denny’s locations per thousand square miles? What about La Quinta?*

```{r dennys states squared miles}
dennys %>%
  count(state) %>%
  inner_join(states, by = c("state" = "abbreviation")) %>%
  mutate(dennys_per_sq= 1000*area/n) %>%
  arrange(dennys_per_sq)
```

The state with the most Denny's per 1000 square miles is Rhode Island, as the District of Columbia is not a state.

```{r laquinta per sq mile}
laquinta %>%
  count(state) %>%
  inner_join(states, by = c("state" = "abbreviation")) %>%
  mutate(laquinta_per_sq= 1000*area/n) %>%
  arrange(laquinta_per_sq)
```

The state with the most La Quinta's per square mile is also Rhode Island.

Next, we put the two datasets together into a single data frame. However before we do so, we need to add an identifier variable. We’ll call this `establishment` and set the value to "Denny's" and "La Quinta" for the `dennys` and `laquinta` data frames, respectively.

```{r identified var establishment}
dennys <- dennys %>%
  mutate(establishment = "Denny's")
laquinta <- laquinta %>%
  mutate(establishment = "La Quinta")
```


Because the two data frames have the same columns, we can easily bind them with the bind_rows function:

```{r combining datasets}
dennys_laquinta <- bind_rows(dennys, laquinta)
```

```{r base code denny/laquinta map}
ggplot(dennys_laquinta, mapping = aes(x = longitude, y = latitude, color = establishment)) +
  geom_point()
```

*11. Filter the data for observations in North Carolina only, and recreate the plot. You should also adjust the transparency of the points, by setting the `alpha` level, so that it’s easier to see the overplotted ones. Visually, does Mitch Hedberg’s joke appear to hold here?*

### Code not working :(

I really wanted to try and subset within my graph, but despite following the guides I found it's not working. So I'm leaving this code to be analyzed by you, to find what's wrong, and going about it a different way below.

```{r dennys_laquinta_NC not working, eval=FALSE, include=TRUE}
ggplot(subset(dennys_laquinta, state %in% "NC", 
      aes(x = longitude, y = latitude, color = establishment))) +
  geom_point() +
  labs(title = "Denny's and La Quinta locations in NC",
       subtitle = "Locations in North Carolina of both chains",
       x = "Longitude", y= "Latitude")
```

### Code that works

```{r working dennys_laquinta NC}
dennys_laquinta_NC <- dennys_laquinta %>%
  filter(state == "NC")
ggplot(dennys_laquinta_NC, 
       aes(x = longitude, y = latitude, color = establishment)) +
  geom_point(alpha=0.5)
```

It does not look to be true that there are frequently Denny's and La Quinta's right next to one another in North Carolina. 

*12. Now filter the data for observations in Texas only, and recreate the plot, with an appropriate ``alpha` level. Visually, does Mitch Hedberg’s joke appear to hold here?*

```{r dennys_laquinta_TX}
dennys_laquinta_TX <- dennys_laquinta %>%
  filter(state == "TX")
ggplot(dennys_laquinta_TX, 
       aes(x = longitude, y = latitude, color = establishment)) +
  geom_point(alpha=0.3)
```

While it is more true in Texas, they are also heavily centered around the state's major cities, so I don't know if I'm convinced. But, it is technically true they hover with one another across the state.